---
title: "session_3"
author: "SMcKnight"
date: "2/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, include = FALSE}
### Install development versions of tmap. https://github.com/mtennekes/tmap

# library(devtools)
# install_github("mtennekes/tmaptools")
# install_github("mtennekes/tmap")
```



### 1. Attach necessary packages

We'll be using **tidiyverse**, **sf**, **tmap** packages! (sf = simple features)


Use control + alt + i to insert a code chunk

```{r, include = FALSE}
#activate packages. 
#using "include = FALSE will make it so this stuff doesn't show up in the knited vesion

library(tidyverse)
library(sf) #Support for simple features, a standardized way to encode spatial vector data. Binds to 'GDAL' for reading and writing data, to 'GEOS' for geometrical operations, and to 'PROJ' for projection conversions and datum transformations.
library(tmap) #With the tmap package, thematic maps can be generated with great flexibility. The syntax for creating plots is similar to that of ggplot2, but tailored to maps

```

### 2. Read in CA eco-regions data
Use the `read_sf()` function to help read all the files of a shapefile. Otherwise you'd have to read each one in individually

```{r}
# dsn = working directory where you will find these files. if you were trying to get files that weren't in your working directory, you'd give it the long tedious path, but since we're already there, you can give it a dot. In my case, i placed the data in a "data" folder so, I put "data"
# layer = the name of the spatial data, of the file prefix name
# When you load something as sf, the geomotry is sticky, i.e., the spatial component of the object are sticky. You can work with all the other attributes as if it were just a normal non-spatial data frame, and R is going to assume you want to retain the spatial information associated with those observations (see teh last 'geometry' field in the attributes)
# If you %>%  and do select(US_L3NAME) it will create a new dataframe with just the US_L3NAME and the geometry will stick to it.
# the rename() function will allow you to rename a field
# st_simplify() will simplify the geomotry. use the dTolerance to adjust the settings. ***You wouldn't want to really do this in a real project, only doing it here to save time*** be very careful about over simplifying. Note that the st_simplify simplifies shapes individually, and doesn't take into account that the shapes should probably be touching. you can use rmshaper::ms_simplify(), geos::gSimplify() to simplify and make sure the features are touching.

ca_eco <- read_sf(dsn = "data", layer = "ca_eco") %>% #read in the data, assign to new variable
  select(US_L3NAME) %>% #return only the US_L3NAME and the sticky geomoetry
  rename(region = US_L3NAME) %>% #rename US_l3NAME to something more meaningful
  st_simplify(dTolerance = 100) %>% #simplify the shape so it draws quicker
  st_transform(crs = 4326) # assign a coordinate reference system of WGS84. 

plot(ca_eco) # plot() is the basic plot function within R

# use st_crs(ca_eco) in the console to verify the coordinate system

```

### 3. Read in CA Counties polygons

```{r}
ca_counties <- read_sf(dsn = "data", layer = "california_county_shape_file")

st_crs(ca_counties) = 4326

```


### 4. Read in CA dams locations & data

```{r}
ca_dams <- read_sf(dsn = "data", layer = "California_Jurisdictional_Dams")

```


### 5. Make a map of dams in California, overlayed on eco-regions, with outlines of CA counties.

```{r}
# ggplot plays really nicely with sf_objects using geom_sf()
# Remember, you always need three things to plot with ggplot(). 1. Tell it we are using ggplot, 2. tell it what data we are using, 3. Tell it what type of geomoetry we want it to produce

#This maps the ca_boundaries first with a black outline and then the eco regions on top of that with a transparency.
# ggplot(data = ca_counties) +
#   geom_sf(color = "black",
#           size = 0.1) +
#   geom_sf(data = ca_eco, #when you want to add a new layer you need to add the data = ca_eco
#           aes(fill = region), #when your color fill depends on a variable in the attributes, it has to be wrapped in the aes()
#           alpha = 0.5, #this will adjust transparency, i.e. 50%
#           color = "NA", #this pertains to the color around each polygon
#           show.legend = TRUE) #if you have FALSE it will not show legend

#this maps the eco regions fist, color coded by region, and then the counties on top of it with hollow fill. This is probably a better way to do it than the top.
ggplot(data = ca_eco) +
  geom_sf(aes(fill = region),
          color = "NA") +
  geom_sf(data = ca_counties,
          fill = "NA") + 
  geom_sf(data = ca_dams,
          size = 0.5,
          alpha = 0.5) +
  theme_minimal()


```

### 6. Just explore SB county eco-regions using intersect

```{r}

sb <- ca_counties %>% 
  filter(NAME == "Santa Barbara")

# clip the eco-region data to only include informatoin ithin SB county, use the st_intersection()

eco_clip <- st_intersection(ca_eco, sb)


# now let's plot it! FIrst using ggplot2:
ggplot() +
  geom_sf(data = ca_counties, 
          fill = "gray90", 
          color = "gray80",
          size = 0.2) + 
  geom_sf(data = eco_clip, 
          aes(fill = region),
          color = "white", 
          size = 0.4) +
  coord_sf(xlim = c(-121, -119), ylim = c(33.5, 35.5)) + 
  scale_fill_manual(values = c("magenta","gold2","blue"))


```

### 7. Create an interactive map with tmap


```{r}

sb_clip_tmap <- tm_basemap("Esri.WorldImagery") + #if you search ?tm_basemap in the console, you should be able to find all the diffent types of basemaps
  tm_shape(eco_clip) + # use the eco_clip layer
  tm_fill("region", #symbolize on region
          palette = c("orange","purple","yellow"),
          alpha = 0.5)

tmap_mode("view") #setting to view mode allows the map to be interactive, default is static
sb_clip_tmap #view the interactive map. (I guess you don't have to use ggplot or plot to view)



```

